---
title: "A summary of the TE landscape of the original dm6 genome"
author: "Ting-Hsuan Chen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: '2'
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    highlight: tango
---

## 0 Setup
### 0.1 Setup genome name and directory
```{r setup, include=FALSE}
genomename <- 'dm6_original'
short_gname <- 'dm6'
setwd('../../')
WRK <- getwd()
SIM <- file.path(WRK, "pub/output/TEgenomeSimulator_dm6_result/")
OUT <- file.path(WRK, "pub/output/sim_report/")
dir.create(file.path(OUT), recursive = TRUE, showWarnings = FALSE)
knitr::opts_knit$set(root.dir=OUT)
```
### 0.2 Load libraries
```{r, load_TEsnoopeR, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
source(file.path(WRK, "pub/script/functions.R"))
```

Tables and images saved in\
`r OUT`

### 0.3 Load files
```{r, load_files, include=FALSE}
sum_file <- read.table(paste0(SIM, 'summarise_repeatmasker_out_family.txt'), header = T, sep = '\t', comment.char = "")
data.sum <- sum_file |> 
  dplyr::mutate(superfamily = gsub("auto", "", superfamily)) |>
    dplyr::mutate(superfamily = gsub("nona", "", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/Helitron", "RC\\/Helitron", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/CMC.*", "DNA\\/CACTA", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/MULE.*", "DNA\\/MULE", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/MuDR.*", "DNA\\/MULE", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/PIF-Harbinger", "DNA\\/Harbinger", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/TcMar.*", "DNA\\/Mariner", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/hAT.*", "DNA\\/hAT", superfamily)) |>
    dplyr::mutate(superfamily = gsub("LTR\\/Gypsy", "LTR\\/Ty3", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA$", "DNA\\/unknown", superfamily)) |>
  group_by(superfamily) |> 
  summarise(count = sum(copynumber), bp = sum(total_length), family_count = n()) |> 
  rename(TE = superfamily) |>
  arrange(TE)

rmout <- read.table(paste0(SIM, 'processed_output.temp'), header = T, sep = '\t', comment.char = "")
rmout <- rmout |> mutate(div = as.numeric(1 - identity), itg = as.numeric(length_ratio))  |>
  dplyr::mutate(superfamily = gsub("auto", "", superfamily)) |>
    dplyr::mutate(superfamily = gsub("nona", "", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/Helitron", "RC\\/Helitron", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/CMC.*", "DNA\\/CACTA", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/MULE.*", "DNA\\/MULE", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/MuDR.*", "DNA\\/MULE", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/PIF-Harbinger", "DNA\\/Harbinger", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/TcMar.*", "DNA\\/Mariner", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA\\/hAT.*", "DNA\\/hAT", superfamily)) |>
    dplyr::mutate(superfamily = gsub("LTR\\/Gypsy", "LTR\\/Ty3", superfamily)) |>
    dplyr::mutate(superfamily = gsub("DNA$", "DNA\\/unknown", superfamily))

rmtbl <- paste0(SIM, "repeatmasker/GCF_000001215.4_Release_6_chromosome_only.fna.tbl")
```

## 1. The proportion of the TE/nonTE sequence
```{r, get_masked_size, echo=FALSE}
lines <- readLines(rmtbl)

# Initialize variables
total_length <- NA
bases_masked <- NA

# Extract values
for (line in lines) {
  if (grepl("total length", line)) {
    genomesize <- gsub(".*total length:", "", line)
    genomesize <- gsub("bp.*", "", genomesize)
    genomesize <- as.integer(genomesize)
  }
  if (grepl("bases masked", line)) {
    bases_masked <- gsub(".*bases masked:", "", line)
    bases_masked <- gsub("bp.*", "", bases_masked)
    bases_masked <- as.integer(bases_masked)
    perc_masked <- gsub(".*\\(", "", line)
    perc_masked <- gsub("%.*", "", perc_masked)
    perc_masked <- as.numeric(perc_masked)
  }
}

# Print extracted values
cat("Total genome size:", genomesize, "bp\n")
cat("Bases masked:", bases_masked, "bp\n")
cat("Percentage masked:", perc_masked, "%\n")

```


```{r, fig1, echo=FALSE, fig.width = 4, out.width="30%"}
plotwidth <- 4
plotheight <- 4
plot <- plot_sim_genome_bp_piechart(genomesize, bases_masked, genomename)
options(repr.plot.width=plotwidth, repr.plot.height=plotheight)
plot
```

```{r, save_fig1, echo=FALSE}
png(filename = paste0(OUT, "genome_size_piegraph_", genomename, ".png"), width = plotwidth, height = plotheight, units = "in", res = 1200)
print(plot)
while (!is.null(dev.list()))  dev.off()
```

File saved as: genome_size_piegraph_`r genomename`.png 

## 2. Breaking down the simulated TEs by superfamily
```{r, table1, echo=FALSE}
total_loci <- sum(data.sum$count)
total_bp <- sum(data.sum$bp)
total_fam <- sum(data.sum$family_count)
data_sum <- data.sum
data_sum$total_loci <- rep(total_loci, nrow(data_sum))
data_sum$total_bp <- rep(total_bp, nrow(data_sum))
data_sum <- data_sum %>% 
  mutate(loci_percentage = round(100*(count/total_loci), 2), bp_percentage = round(100*(bp/total_bp), 2), family_percentage = round(100*(family_count/total_fam), 2)) %>%
  rename(TE_superfamily = TE, loci = count) %>%
  select(-total_loci, -total_bp)
total_row <- data.frame(TE_superfamily = "Total", loci = total_loci, bp = total_bp, family_count = total_fam, loci_percentage = 100, bp_percentage = 100, family_percentage = 100)
data_sum <- rbind(data_sum, total_row)
write.table(data_sum, paste0(OUT, "genome_summary_TEsuperfamily_", genomename, ".csv"), col.names = T, row.names = F, sep = ",")
kable(data_sum, format="markdown")
```

Saved file name: genome_summary_TEsuperfamily_`r genomename`.csv 

### 2.1 Total TE loci categorised by TE superfamily

```{r, fig2, echo=FALSE, out.width="60%"}
plotwidth <- 8
plotheight <- 8
options(repr.plot.width=plotwidth, repr.plot.height=plotheight)
plot <- plot_te_loci_piechart(data.sum, genomename)
plot
```

```{r, save_fig2, echo=FALSE}
png(filename = paste0(OUT, "te_loci_piegraph_", genomename, ".png"), width = plotwidth, height = plotheight, units = "in", res = 1200)
print(plot)
while (!is.null(dev.list()))  dev.off()
```
High quality image saved in\
`r OUT`

Saved file name: te_loci_piegraph_`r genomename`.png 

### 2.2 Total simulated TE bases categorised by TE superfamily


```{r, fig3, echo=FALSE, out.width="60%"}
plotwidth <- 8
plotheight <- 8
options(repr.plot.width=plotwidth, repr.plot.height=plotheight)
plot <- plot_te_bp_piechart(data.sum, genomename)
plot
```

```{r, save_fig3, echo=FALSE}
png(filename = paste0(OUT, "te_bp_piegraph_", genomename, ".png"), width = plotwidth, height = plotheight, units = "in", res = 1200)
print(plot)
while (!is.null(dev.list()))  dev.off()
```

Saved file name: te_bp_piegraph_`r genomename`.png 

### 2.3 The distribution of TE loci divergence 


```{r, fig9, echo=FALSE, fig.width = 8, fig.asp = .5, out.width="60%"}
plotwidth <- 8
plotheight <- 4
options(repr.plot.width=plotwidth, repr.plot.height=plotheight)
ymax <- 2500
ybreak <- 500
plot <- plot_ori_divergence_by_loci(rmout, ymax, ybreak, genomename)
plot
```

```{r, save_fig9, echo=FALSE}
png(filename = paste0(OUT, "te_loci_divergence_", genomename, ".png"), width = plotwidth, height = plotheight, units = "in", res = 1200)
print(plot)
while (!is.null(dev.list()))  dev.off()
```

Saved file name: te_loci_divergence_`r genomename`.png 


### 2.4 The distribution of TE loci integrity

```{r, fig10, echo=FALSE, fig.width = 8, fig.asp = .5}
plotwidth <- 8
plotheight <- 4
options(repr.plot.width=plotwidth, repr.plot.height=plotheight)
ymax <- 4000
ybreak <- 500
plot <- plot_ori_integrity_by_loci(rmout |> filter(itg >= 0 & itg <= 1), ymax, ybreak, genomename)
plot
```

```{r, save_fig10, echo=FALSE}
png(filename = paste0(OUT, "te_loci_integrity_", genomename, ".png"), width = plotwidth, height = plotheight, units = "in", res = 1200)
print(plot)
while (!is.null(dev.list()))  dev.off()
```

Saved file name: te_loci_integrity_`r genomename`.png 
